Type Safe Redis Queries

### 2016-06-17 aplas16/main-aplas.tex
Section 2 done.
Half of Section 3.3.
Revision up to incr and incrfloat.
Moved some examples to Introduction.
Up to assertions, and the final example.
Another round of revision from the beginning. Corrected some bibliography.

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{color}
\usepackage{url}
\usepackage{mathptmx}
\usepackage{doubleequals}
\usepackage{makeidx}  % allows for indexgeneration

@@

\title{Safe Redis Queries}
\subtitle{A Case Study of Type-Level Programming}
\author{Ting-Yan Lai \inst{1}\inst{2}\and Tyng-Ruey Chuang \inst{1}
  \and Shin-Cheng Mu \inst{1}}
\institute{Institute of Information Science, Academia Sinica, Taiwan
\and Institute of Computer Science and Engineering,
National Chiao Tung University, Taiwan}

@@

\begin{abstract}
\Redis{} is an in-memory data strcture store, often used as a database, with a
Haskell interface \Hedis{}. \Redis{} is dynamically typed --- a value associated
with a key can be overwritten by a value of a different type, and a command,
when fetching a value of a type it does not expect, signals a runtime error. We
develop a domain-specific language that, by exploiting Haskell type-level
programming techniques including indexed monad, type-level literals and closed
type families, keeps track of types of values in the database and statically
guarantees that type errors cannot happen.
\end{abstract}

\input{sections/Introduction}
%\input{sections/Motivation}
\input{sections/TypeLevelDict}
\input{sections/IndexedMonads}
\input{sections/TypeLevelFun}
\input{sections/PolyRedis}
\input{sections/Constraints}
\input{sections/Assertions}
\input{sections/Discussions}
\input{sections/Related}
\input{sections/Conclusions}

>>>>

\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{color}
\usepackage{url}
\usepackage{mathptmx}
\usepackage{doubleequals}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{textcomp} % for single quote
\usepackage{fancyvrb} % for indented verbatim

@@

\title{Type Safe Redis Queries}
\subtitle{A Case Study of Type-Level Programming in Haskell}
\author{Ting-Yan Lai \inst{1}\inst{2}\and Tyng-Ruey Chuang \inst{1}
  \and Shin-Cheng Mu \inst{1}}
\institute{Institute of Information Science, Academia Sinica, Taiwan
\and Institute of Computer Science and Engineering,
National Chiao Tung University, Taiwan}

@@

\begin{abstract}
\Redis{} is an in-memory data structure store, often used as a database, with a
Haskell interface \Hedis{}. \Redis{} is dynamically typed --- a key can be
discarded and re-associated to a value of a different type, and a command,
when fetching a value of a type it does not expect, signals a runtime error. We
develop a domain-specific language that, by exploiting Haskell type-level
programming techniques including indexed monad, type-level literals and closed
type families, keeps track of types of values in the database and statically
guarantees that type errors cannot happen for a class of \Redis{} programs.
\end{abstract}

\input{sections/Introduction}
\input{sections/IndexedMonads}
\input{sections/TypeLevelDict}
\input{sections/Embedding}
\input{sections/Discussions}
\input{sections/Conclusions}

### 2016-06-17 aplas16/sections/Conclusions.tex
Another round of revision from the beginning. Corrected some bibliography.

\section{Conclusions}
\label{sec:conclusions}

We have demonstrated how a DSL with variable bindings could be embedded in types.
With more and more extensions added to the language, Haskell is gradually
 becoming a dependently-typed language,\footnote{Why not be dependently typed? \url{http://stackoverflow.com/questions/12961651/why-not-be-dependently-typed}}
 but it's not that dreadful as many (including us) would have thought.

>>>>

\section{Conclusions and Related Work}
\label{sec:conclusions}

By exploiting various recent extensions and type-level programming techniques,
we have designed a domain-specific embedded language \Edis{} which enforces
typing disciplines that keeps track of available keys and their types. The
type of a command clearly specifies which keys must or must not present in
the data store, what their types ought to be, as well as how the keys and types are updated after the execution. A program can be constructed only if it does
not throw a runtime type error, and the type also serves as documentation of the
commands. We believe that it is a neat case study of application of type-level programming.

\Redis{} identifies itself as a data structure store/server, rather than a
database. Nevertheless, there has been attempts to design DSELs for relational
databases that guarantee all queries made are safe. Among them, {\sc HaskellDB}~\cite{haskelldb,haskelldbimproved} dynamically generates, from monad comprehensions, SQL queries to be executed on ODBC database servers.
With the expressiveness of dependent types, Oury and Swierstra~\cite{pi}
build a DSEL in Agda for relational algebra. After GHC introduced more features
faciliting type-level programming, Eisenberg and Weirich~\cite{singletons}
ported these techniques to Haskell using singleton types.

None of the type-level programming techniques we used in this paper are
particularly new. Indexed monads (also called parameterized monads) have been
introduced by Atkey~\cite{indexedmonad}.
Kiselyov et al~\cite{typefun} used indexed monads to track the locks held among
a given finite set. The same paper also demonstrated implementation of a
variety of features including memoization, generic maps, session types, typed \texttt{printf} and \texttt{sprintf}, etc., by type-level programming.
Before the introduction of type families, Kiselyov and Shan~\cite{staticresources} used type classes and functional dependencies to
implement type-level functions, and showed that they are sufficient to track
resources in device drivers.

\cite{constraintskinds}

Lindley and McBride~\cite{phasedistinction} provided a good tutorial, analysis, and summary of the dependent type like features currently in Haskell, and
compared them with ``real'' dependently typed languages.
